Exploring probability distributions for bivariate temporal granularities

### 2019-12-04 main.Rmd
demystifying definitions using day and month example

\newtheorem{timedom}{Definition}
\begin{timedom}\label{def:timedom}
A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
\end{timedom}

@@

**Discussion**:  In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorisation of the circular granularity. However, the labels can coincide with indexes in which case integers are directly used to refer to categorisations of the circular granularity.  Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}. Each granule is represented by a box. The diagram also illustrates that the granules that overlap share elements from the underlying time domain. The first slot in the diagram shows the index set and the bottom granularity. The index of the tsibble considered is days. G and H are two linear granularities denoting days and weeks respectively. $k(day, week) =  7$ and hence, $C_{G, B}(z)$ represented by C(BG, G) is given by $B(z\mod 7)$. The circular granularity C(B, G) consists of repetitive pattern {B(0), B(1), B(2), B(3), B(4), B(5), B(6)} which repeats itself after each period length. Similarly, C(BG, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14. C(B, H) also repeats the pattern ${B(0), B(1), \dots, B(13)}$ in every period. Hence, circular granularities are represented as repetitive categorization of time in the diagram.

@@

In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = BG(\lfloor z/k(BG,G) \rfloor\mod k(G,H))$, where linear granularity H is periodic with respect to linear granularity G and k(G, H) represents the period length of G within H.

Table \ref{tab:definitions} shows representation of circular granularities relating two linear granularities. It is possible that none of these two linear granularities are bottom granularities. But the representation of the resultant circular granularity will be a function of the bottom granularity. Suppose the bottom granularity is minutes, and $k_i$ is the number of granules of the finer granularity in each granule of the coarser granularity in the circular granularity $C_i$, also referred to as the period length.

\begin{table}[ht]
\begin{center}
\begin{tabular}{lr@{~}lr@{~}r}
\toprule
Minute-of-Hour: & $C_1(z)$ & $= z \mod 60$ & $k_1 =$&$60$ \\
Minute-of-Day: & $C_2(z)$ & $= z \mod 60*24$ & $k_2=$&$`r 60*24`$\\
Hour-of-Day: & $C_4(z)$ & $= \lfloor z/60\rfloor\mod 24$  & $k_3 =$&$24$ \\
Hour-of-Week: & $C_6(z)$ & $= \lfloor z/60\rfloor\mod 24*7$  & $k_4=$&$`r 24*7`$\\
Day-of-Week: & $C_7(z)$ & $= \lfloor z/24*60\rfloor \mod 7$ & $k_5=$&$7$\\
\bottomrule
\end{tabular}
\end{center}
\caption{Illustrative circular granularities with time index in minutes}
\label{tab:definitions}
\end{table}

@@

A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity. For example, \autoref{tab:quasi} shows some example of quasi-circular granularities ($QC_i$) and plausible choices of number of granules ($k_i$) of the finer granularity inside each granule of the coarser one in the cyclic granularity ($QC_i$).

@@

Day-of-Month: & $QC_1(z)$ && $k_1=31, 30, 29, 28$\\
Hour-of-Month: & $QC_2(z)$ && $k_2=24*31, 24*30, 24*29, 24*28$\\
Day-of-Year: & $QC_3(z)$ && $k_3=366, 365$\\
Week-of-Month: & $QC_4(z)$ && $k_4=5, 4$\\ 
\bottomrule
\end{tabular}
\end{center}
\caption{Illustrative quasi-circular granularities with potential period lengths}
\label{tab:quasi}
\end{table}

\newtheorem{timedom1}{Definition}[subsection]
\begin{timedom1}\label{def:quasicircular}
A quasi-circular granularity QC_(BG, G') that relates linear granularities G' and bottom granularity BG, if

\begin{equation}\label{eq:eq3}
QC_{BG, G'}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
\end{equation}

where,  
z denotes the index set,  
BG denotes the index of the tsibble (bottom granularity),  
G' is a linear granularity periodic with respect to BG with irregular mapping,   
R' denotes the number of granules of G' in each period of the grouping (BG,G),   
T_w are the sets of indices of BG describing G'(w) such that $G'(w) = \bigcup_{z \in T_w}BG(z)$,$T_w = 0$ for $w \epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
\end{timedom1}

@@

**Discussion**: Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-gran}. Each granule is represented by a box. Similar to \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity. 
Two linear granularities AG and AH are considered, where $G' = Alter^{2}_{(1,-1)}(BG, Group_3(BG))$ and $H' = Alter^{2}_{(1,-2)}(BG, Group_7(BG))$. This implies that G' is made up by shrinking every 1st granule of Group_3(BG) by 1 granule and H' is made up of shrinking every 1st granule of Group_3(BG) by 2 granules. Number of granules of G' and H' in each period of BG is 2, but the number of granules of BG in each of those granules are different. QC(BG, G') and QC(BG, H') are repetitive categorization of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of G' or H'. For G', R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}. For H', R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.

@@

\begin{timedom}\label{def:acyclic}
An aperiodic granularity A_(BG, G'') that relates linear granularities G and bottom granularity BG, with G'' being aperiodical with respect to BG, is given by

>>>>

\newtheorem{timedom}{Definition}[section]

\begin{timedom}\label{def:timedom}
A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
\end{timedom}

@@

**Discussion**:  In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorisation of the circular granularity. However, the labels can coincide with indexes in which case integers are directly used to refer to categorisations of the circular granularity. Note that each
circular granularity can use a different set of labels. Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}. Each granule is represented by a box. The diagram also illustrates that the granules that overlap share elements from the underlying time domain. The first slot in the diagram shows the index set and the bottom granularity. The index of the tsibble considered is days. G and H are two linear granularities denoting days and weeks respectively. $k(day, week) =  7$ and hence, $C_{G, B}(z)$ represented by C(BG, G) is given by $B(z\mod 7)$. The circular granularity C(B, G) consists of repetitive pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length. $L$ can be defined such that the labels corresponding to ${0,1,\dots,6}$ correspond to different days of the week ${Sun, Mon ,\dots, Sat}$ . Similarly, C(B, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14. C(B, H) also repeats the pattern ${L'(0), \dots, L'(13)}$ in every period. Hence, circular granularities are represented as repetitive categorization of time in the diagram.

@@

In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/k(B,G) \rfloor\mod k(G,H))$, where linear granularity H is periodic with respect to linear granularity G and k(G, H) represents the number of granules of G in one granule of H.

Table \ref{tab:definitions} shows representation of circular granularities relating two linear granularities. It is possible that none of these two linear granularities are bottom granularities. But the representation of the resultant circular granularity will be a function of the index set. Suppose the bottom granularity is minutes, $k_i$ is the number of granules of the finer granularity in each granule of the coarser granularity in the circular granularity $C_i$ and $L_i$ is the associated label mapping.

\begin{table}[ht]
\begin{center}
\begin{tabular}{lr@{~}lr@{~}r}
\toprule
Minute-of-Hour: & $C_1(z)$ & $= L_1(z \mod 60)$ & $k_1 =$&$60$ \\
Minute-of-Day: & $C_2(z)$ & $= L_2(z \mod 60*24)$ & $k_2=$&$`r 60*24`$\\
Hour-of-Day: & $C_4(z)$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $k_3 =$&$24$ \\
Hour-of-Week: & $C_6(z)$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $k_4=$&$`r 24*7`$\\
Day-of-Week: & $C_7(z)$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $k_5=$&$7$\\
\bottomrule
\end{tabular}
\end{center}
\caption{Illustrative circular granularities with time index in minutes}
\label{tab:definitions}
\end{table}

@@

A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity. For example, \autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) and plausible choices of number of granules ($k_i$) of the finer granularity inside each granule of the coarser one in the cyclic granularity ($Q_i$).

@@

Day-of-Month: & $Q_1(z)$ && $k_1=31, 30, 29, 28$\\
Hour-of-Month: & $Q_2(z)$ && $k_2=24*31, 24*30, 24*29, 24*28$\\
Day-of-Year: & $Q_3(z)$ && $k_3=366, 365$\\
Week-of-Month: & $Q_4(z)$ && $k_4=5, 4$\\ 
\bottomrule
\end{tabular}
\end{center}
\caption{Illustrative quasi-circular granularities with potential period lengths}
\label{tab:quasi}
\end{table}

\begin{timedom}\label{quasicircular}
A quasi-circular granularity $Q_(B, G')$ that relates linear granularities $G'$ and bottom granularity $B$, if

\begin{equation}\label{eq:eq3}
Q_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), \quad z \in T_k
\end{equation}

where,  
$z$ denotes the index set,\\
$B$ denotes a full-integer labelled bottom granularity which groups periodicially into linear granularity $G'$ with irregular mapping,\\
{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,\\
$T_w$ are the sets of indices of B describing $G'(w)$ such that 
\begin{equation}
G'(w) = \bigcup_{z \in T_w}B(z),\\
\end{equation}
$T_w = 0$ for $w \epsilon Z_{<0}$,\\
$\vert T_w \vert$ is the cardinality of set $T_w$.
\end{timedom}

@@

**Discussion**: According to the first condition, assume, for instance, that G'(w) is the union of the granules G(a1), G(a2), . . . , G(ak). The periodicity property (condition (2) in the definition) ensures that the Rth granule after H(i), i.e., H(i + R), if non-empty, is the union of G(a1 + P ), G(a2 + P), . . . , G(ak + P).

Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-gran}. Each granule is represented by a box. Similar to \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity. 
Two linear granularities AG and AH are considered, where $G' = Alter^{2}_{(1,-1)}(BG, Group_3(BG))$ and $H' = Alter^{2}_{(1,-2)}(BG, Group_7(BG))$. This implies that G' is made up by shrinking every 1st granule of Group_3(BG) by 1 granule and H' is made up of shrinking every 1st granule of Group_3(BG) by 2 granules. Number of granules of G' and H' in each period of BG is 2, but the number of granules of BG in each of those granules are different. QC(BG, G') and QC(BG, H') are repetitive categorization of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of G' or H'. For G', R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}. For H', R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.

@@

\begin{timedom}\label{aperiodic}
An aperiodic granularity A_(BG, G'') that relates linear granularities G and bottom granularity BG, with G'' being aperiodical with respect to BG, is given by
